---
title: "Mel Schwan DDS Unit 5 HW"
author: "Mel Schwan"
date: "5/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MSDS 6306: Doing Data Science 
#Live session Unit 5 assignment
Backstory: Your client is expecting a baby soon.  However, he is not sure what to name the child.  Being out of the loop, he hires you to help him figure out popular names.  He provides for you raw data in order to help you make a decision

# Question 1 COPY 
	1.	Data Munging (30 points): Utilize yob2016.txt for this question. This file is a series of popular children’s names born in the year 2016 in the United States.  It consists of three columns with a first name, a gender, and the amount of children given that name.  However, the data is raw and will need cleaning to make it tidy and usable.
a.	First, import the .txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you’re dealing with.  Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.
 
	
	
```{r}

#Question 1 part a.
child_name_df <- read.delim2(file="yob2016.txt", sep=';', head=FALSE)
head(child_name_df,10)
names(child_name_df) <- c('Name', 'Sex', 'Count')
head(child_name_df,10)

#Question 1 b.	Display the summary and structure of df
summary.data.frame(child_name_df)
dim.data.frame(child_name_df)

#Question 1 c.
#Find row with the yyy
row = grep("yyy$", child_name_df[,1])
#Question 1 d
#Once row is identified then remove the row and reassign to same dataframe
child_name_df <- child_name_df[-c(212),]
row = grep("yyy", child_name_df[,1])

```
2.	Data Merging (30 points): Utilize yob2015.txt for this question.  This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.

a.	Create a new column called “Total” in final that adds the amount of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?


```{r}

child_name_df2 <- read.delim2(file="yob2015.txt", sep=',', head=FALSE)
names(child_name_df2) <- c('Name', 'Sex', 'Count')
#Print last 10 records
tail(child_name_df,10)
#Observation they are all males and the count is equal for all ten names.

```
c.	Merge y2016 and y2015 by your Name column; assign it to final.  The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.

```{r}

final <- merge.data.frame(child_name_df, child_name_df2, by.x = "Name", by.y = "Name", fall.x =FALSE, sort = TRUE )


```

3.	Data Summary (30 points): Utilize your data frame object final for this part.

a.	Create a new column called “Total” in final that adds the amount of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?

```{r}
final$Total <- 0
head(as.numeric(final$Count.x, final$Count.y))
final$Total=final$Count.x+final$Count.y
finalsum <- sum(final$Total)
finalsum
#final <- order(final$Total)
#finalorder




```